---
title: "STAT 4410/8416 Homework 3"
author: "lastName firstName"
date: "Due on Tuesday Nov 6, 2022"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(fig.align='center', message=FALSE, warning=FALSE)
output = opts_knit$get("rmarkdown.pandoc.to")
if(!is.null(output)) {
  if(output=="html") opts_chunk$set(out.width = '400px') else
    opts_chunk$set(out.width='.6\\linewidth')
}
```

**1.** **Text Data analysis:** Download `lincoln-last-speech.txt` from Canvas which contains Lincoln's last public address. Now answer the following questions and include your codes.  

   a)  Read the text and store the text in `lAddress`. Show the first 70 characters from the first element of the text.  

```{r}
library(stringr)
lAddress <- readLines("C:/Users/Nelson Lefebvre/Desktop/AUTRE/lincoln-last-speech.txt")
print(substr(lAddress[1], start=0, stop=70))
```

   
   b)  Now we are interested in the words used in his speech. Extract all the words from `lAddress`, convert all of them to lower case and store the result in `vWord`. Display first few words.  
   
```{r}
lAddress <- paste(lAddress, collapse="")
vWord <- tolower(lAddress)
print(substr(vWord[1], start=0, stop=70))
```
   
   
   c)  The words like `am`, `is`, `my` or `through` are not much of our interest and these types of words are called stop-words. The package `tm` has a function called `stopwords()`. Get all the English stop words and store them in `sWord`. Display few stop words in your report.  
   
```{r}
#install.packages("stopwords")
#install.packages("tm")
library("stopwords")
library("tm")
sword = stopwords(kind = "en")
# source : https://cran.r-project.org/web/packages/stopwords/stopwords.pdf
print(sword)
```
   
   
   d) Remove all the `sWord` from `vWord` and store the result in `cleanWord`. Display first few clean words.  

```{r}
temp <- lAddress
temp <- gsub('[^[:alnum:] ]','',temp)
cleanWord <- unlist(strsplit(temp, " "))
cleanWord <- cleanWord[!cleanWord %in% sword]
cleanWord <- cleanWord[!duplicated(cleanWord)]
print(cleanWord)
```

   
   e)  `cleanWord` contains all the cleaned words used in Lincoln's address. We would like to see which words are more frequently used. Find 15 most frequently used clean words and store the result in `fWord`. Display first 5 words from `fWord` along with their frequencies.  

```{r}
library("stringr")
library(tidyr)
library(dplyr)
fWord <- data.frame(cleanWord,str_count(lAddress, cleanWord))
colnames(fWord) <- c('C1','C2')
fWord <- fWord %>% 
 # desc orders from largest to smallest
 arrange(desc(C2)) %>%
 # select subsets the columns you want
 select(C1, C2)
fWord <- data.frame(fWord$C1[1:15],fWord$C2[1:15])
colnames(fWord) <- c('C1','C2')
print(fWord)
head(fWord, 5)


```
   
   
   f)  Construct a bar chart showing the count of each words for the 15 most frequently used words. Add a layer `+coord_flip()` with your plot.  
```{r}
library(ggplot2)
Db1 <- fWord$C1
Db2 <- fWord$C2
hp <- data.frame(Db1,Db2)
# Horizontal histogram
ggplot(hp, aes(x=Db1, y=Db2)) + 
  geom_bar(stat = "identity") + coord_flip()
```
   
   
   g)  What is the reason for adding a layer `+coord_flip()` with the plot in question (f). Explain what would happen if we would not have done that.  
   
```{r}
ggplot(hp, aes(x=Db1, y=Db2)) + 
  geom_bar(stat = "identity")
```
The problem will be that the name will be superposed because they are long.
   
   h)  The plot in question (f) uses bar plot to display the data. Can you think of another plot that delivers the same information but looks much simpler? Demonstrate your answer by generating such a plot.   
   
```{r}
library(wordcloud)
wordcloud(words = hp$Db1, freq = hp$Db2, min.freq = 1,
          max.words=15, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(15, "Dark2"))
```
   

   i) In the question (c), you removed words that are called `stop-words`. Now please answer the following:
         a)  Count the total stop words from `lAddress` and store it in `stopWordsCount`
```{r}
stopWordsCount <- data.frame(sword,str_count(lAddress, sword))
stopWordsCount[stopWordsCount==0] <- NA
stopWordsCount <- stopWordsCount[complete.cases(stopWordsCount),]
colnames(stopWordsCount) <- c('C1','C2')
print(stopWordsCount)
```
         
         b)  Count the total words (including stop-words) from `lAddress` and store it in `lAddressCount`
```{r}
lAddressCount <- lengths(gregexpr("\\W+", lAddress)) + 1
print(lAddressCount)
```
         
         c)  Divide `stopWordsCount` by `lAddressCount` and report the percentage
```{r}
stopWordsCount
stopWordsCount$C2 <- stopWordsCount$C2 / lAddressCount
print(stopWordsCount)
#If we want it as a classical pourcentage
stopWordsCount$C2 <- stopWordsCount$C2 * 100
print(stopWordsCount)
```
         
         d)  Explain in your own words what does the percentage indicate in this context?
I think in that case it's showing that the narrator is using to much of one stop word and maybe he should try to take another one when he can, to avoid redundancy.


**2.** **Regular Expressions:** Write a regular expression to match patterns in the following strings. Demonstrate that your regular expression indeed matched that pattern by including codes and results. Carefully review how the first problem is solved for you. 

   aa) We have a vector `vText` as follows. Write a regular expression that matches `g, og, go or ogo` in `vText` and replace the matches with '.'.
```{r}
vText <- c('google','logo','dig', 'blog', 'boogie' )
```

**Answer:**
```{r}
pattern <- 'o?go?'
gsub(pattern, '.', vText)
```
   a)  We have a vector `VNumbers` as follows. Write a regular expression that extracts only binary values (0,1) from `VNumbers`.
```{r}
      VNumbers = c('01011123AEX','1010183DIS','1A02L01A13', 'A2NE3000111')
a <- gsub('[2-9]*[A-Z]', '', VNumbers)
print(a)
```
   b)  Replace only the 5 or 6 digit numbers with the word "found" in the following vector. Please make sure that 3, 4, or 7 digit numbers do not get changed.
```{r}
#i dont know how i should understand the question, so there is two way i done this one
      vPhone = c('874','6783','345345', '32120', '468349', '8149674' )
b <- gsub('[5-6]', 'found', vPhone)
print(b)
b <- gsub('^\\d{5,6}$', 'found', vPhone)
print(b)
```  
   c)  Replace all the characters that are not among the 26 English characters or a space. Please replace with an empty spring.
```{r}
myText = "#y%o$u @g!o*t t9h(e) so#lu!tio$n c%or_r+e%ct"
c <- gsub('[#%$@*!+_9()]', '', myText)
print(c)
```  
   d)  Extract all the Proper nouns from the `myText` using Regular expression.
```{r}
myText <- c
library(stringr)
myText = "welcome to University of Nebraska Omaha. For Math department classes please go to DSC"
d <- str_extract_all(myText, '(?<![:punct:] )[:upper:][:alpha:]*')
print(d)
```  
   e)  Extract all the three numbers embedded in the following text.
```{r}
bigText = 'There are four 20@20 numbers hid989den in the 500 texts'
gsub('[0-9]', '', bigText)
e <- str_extract_all(bigText, '[0-9]+')
print(e)
```  
   f)  Extract all the words all the words that have a upper case letter in the start and end,convert all the words into lowercase.
```{r}
      myText = 'ThE SalrieS are ReporteD (in millions) for every CompanY.'
lapply(str_extract_all(myText, '[:upper:][:alpha:]*[:upper:]'), tolower)
```  
   g)  Extract the texts in between _ and dot(.) in the following vector. Your output should be 'bill', 'pay', 'fine-book'.  
```{r}
      myText = c("H_bill.xls", "Big_H_pay.xls", "Use_case_fine-book.pdf")
myText <- c(gsub('.+\\_{1}', '', myText))
myText <- gsub('\\.{1}.+','', myText)
print(myText)
```  
   h)  Extract the numbers (return only integers) that are followed by the units 'ml' or 'lb' in the following text.   
```{r}
      myText = 'Received 10 apples with 200ml water at 8pm with 15 lb meat and 2lb salt'
lb = str_extract_all(myText, '(?<=( ))[[:digit:]]*( )?(?=(lb))')
ml = str_extract_all(myText, '(?<=( ))[[:digit:]]*(?=(ml))')
print(lb)
print(ml)
```  
   i)  Extract only the word in between pair of symbols `$`. Count number of words you have found between pairs of dollar sign `$`.  
```{r}
      myText = 'Math symbols are $written$ in $between$ dollar $signs$'
str_extract_all(myText, '(?<=(\\$))[[:alpha:]]*(?=(\\$))')
```  
   j)  Extract all the valid equations in the following text.
```{r}
      myText = 'equation1: 21 - 12 = 9, equation2 is: 2*3=6, do not extract 2w3=6'
str_extract_all(myText, '(?<=( ))[^[:alpha:]]*(?=(,))')
```  
   k)  Extract all the letters of the following sentence and check if it contains all 26 letters in the alphabet. If not, produce code that will return the total number of unique letters that are included and show the letters that are missing.
```{r}
      myText = 'there are five wizard boxing matches to be judged'
news=str_extract_all(myText, '[:alpha:]')
sort_news=str_sort(news[[1]], decreasing = FALSE)
un=unique(sort_news)
lun=list(un)
length(lun[[1]])
setdiff(letters,un)
```
   l) Extract the valid web link from the text
```{r}
      myText = '<body> the valid site is http://www.bbc.com not http://.com'
str_extract_all(myText, '(?=(http:)).+(.com )')
```
   

**3.** **Extracting data from the web:** Our plan is to extract data from web sources. This includes email addresses, phone numbers or other useful data. The function `readLines()` is very useful for this purpose.  

   a)  Read all the text in https://www.unomaha.edu/college-of-arts-and-sciences/mathematics/about-us/directory/index.php and store your texts in `myText`. Show first few rows of `myText` and examine the structure of the data.  

```{r}
myText <- readLines('https://www.unomaha.edu/college-of-arts-and-sciences/mathematics/about-us/directory/index.php')
head(myText)
```

   
   b)  Write a regular expression that would extract all the http web links addresses from `myText`. Include your codes and display the results that show only the `http` or `https` web link addresses and nothing else.
   
```{r}
unlist(regmatches(myText, gregexpr("https?...+[_a-z0-9-]+(\\.[a-z]{3,4})", myText)))
```

   
   c)  Now write a regular expression that would extract all the emails from `myText`. Include your codes and display the results that show only the email addresses and nothing else. 

```{r}
unlist(regmatches(myText, gregexpr("([_a-z0-9-]+(\\.[_a-z0-9-]+)*@[a-z0-9-]+(\\.[a-z0-9-]+)*(\\.[a-z]{2,4}))", myText)))
```

   
   d)  Write a regular expression to extract words with 11 or more letters in the text. Include your codes and display the result that shows the words without duplication.

```{r}
unique(unlist(regmatches(myText, gregexpr("\\b[a-zA-Z]{11,}\\b", myText))))
```

   
   e)  Now we want to extract all the phone/fax numbers in `myText`. Write a regular expression that would do this. Demonstrate your codes showing the results without duplication. 
   
```{r}
unique(unlist(regmatches(myText, gregexpr("[0-9]{3}\\.[0-9]{3}\\.[0-9]{4}", myText))))
```

   
   f)  The link of ggplot2 documentation is https://ggplot2-book.org/individual-geoms.html and we would like to get the list of individual ggplot2 geoms from there. Write a regular expression that would extract all the geoms names (geom_bar is one of them) from this link and display the unique geoms. How many unique geoms does the page list?  

```{r}
linked <- readLines('https://ggplot2-book.org/individual-geoms.html')
unique(unlist(regmatches(linked, gregexpr("geom\\_[a-z]{3,7}", linked))))
```
   

**4.** **Big data problem:** Download the sample of a big data `bigDataSample.csv` from Canvas. Note that the data is in csv format and compressed for easy handling. You may need to uncompress it before using. Now answer the following questions. 

   a)  Read the data and select only the columns that contains the word 'human'. Store the data in an object `dat`. Report first few rows of your data.  
   
```{r}
library(dplyr)
dat<- read.csv(file = 'C:/Users/Nelson Lefebvre/Desktop/AUTRE/bigDataSample.csv')
dat<-dat%>%select(contains("human"))
head(dat)
```
   
   b)  The data frame `dat` should have 5 columns. Rename the column names keeping only the last character of the column names. So each column name will have only one character. Report first few rows of your data now.  
   
```{r}
names(dat) <- str_sub(names(dat),-1)
head(dat)
```
   
   
   c)  Compute and report the means of each columns group by column b in a nice table.
   
```{r}
data <-dat %>% group_by(b) %>% summarise_all("mean")
kable(data)
```

   
   d)  Change the data into long form using id='b' and store the data in `mdat`. Report first few rows of data.  
   
```{r}
library(tidyr)
mdat<-gather_(dat, "variable", "value", c("g", "p", "e", "n"))
print(mdat)
```

   
   e)  The data frame `mdat` is now ready for plotting. Generate density plots of value, color and fill by variable and facet by b.  
   
```{r}
library(ggplot2)
mdat%>%mutate_at(vars(b),list(factor))%>%
ggplot(aes(x=value,color=variable,fill=variable))+ geom_density()+facet_wrap(~b)
```

   
   f)  This data set is a sample of much bigger data set. Here we read the data set and then selected the desired column. Do you think it would be wise do the same thing with the actual larger data set? Explain how you will solve this problem of selecting few columns (as we did in question a) without reading the whole data set first. Demonstrate that showing your codes. 
   
```{r}
library(readr)
plop <- read_csv('C:/Users/Nelson Lefebvre/Desktop/AUTRE/bigDataSample.csv',col_select = contains("human"))
print(plop)
```

