---
title: "STAT 4410/8416 Homework 5"
author: "lastName firstName"
date: "Due on Nov 30, 2022"
output:
  pdf_document: 
    latex_engine: xelatex
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(fig.align='center', dpi=100, message=FALSE, warning=FALSE, cache=TRUE)
output <- opts_knit$get("rmarkdown.pandoc.to")
if(!is.null(output)) {
  if (output=="html") opts_chunk$set(out.width = '400px') else
    opts_chunk$set(out.width='.6\\linewidth')
}
```

1. **Working with databases:** Please follow the instruction below before answering the questions:

* Install the package `sqldf` using `install.packages('sqldf')`
* Import the library using `library('sqldf')`
* Read the file `https://raw.githubusercontent.com/dsindy/kaggle-titanic/master/data/train.csv ` and store it in an object called `titanic`

```{r, echo=FALSE}
library(data.table)
titanic <- fread('https://raw.githubusercontent.com/dsindy/kaggle-titanic/master/data/train.csv')
```


We can now start writing SQL Script using `SQLDF library` right inside R. See example below:

```{r}
library(sqldf)

sqldf("SELECT passengerid, name, sex
      FROM titanic
      limit 5", drv="SQLite")
```


Answer the following questions. Write SQL Script where applicable.

a. What does the following command do in MySQL?
    i. `show databases;`
    It's showing a list of all the all the databases
    ii. `show tables;`
    It's showing a list of all the all the tables
b. Write `SQL` script to answer the following questions based on titanic data. Display the results of your script.
    i. What is the average age of passengers who survived? Group the data by Sex. Display only the column `Sex`, `AverageAge`
```{r}
sqldf("SELECT sex, avg(age) AS avg_age
FROM titanic
WHERE survived=1
GROUP BY sex", drv='SQLite')
```
    ii. What is the percentage of passengers who survived in each Passenger Class or `Pclass`? Group the data by `Sex`. Display Pclass, Sex, percentage value. 
```{r}
sqldf("SELECT pclass, sex,
COUNT(survived=1)/COUNT(*) OVER (PARTITION BY sex) AS Percent
FROM titanic
GROUP BY pclass, sex", drv='SQLite')
```
    iii. What is the average age of all the passenger (survived and not survived)? Group the data by `Pclass`, `Sex`, `Survived`. After that use `ggplot` to generate a line plot to show average age vs pclass, facet by sex and color it by survived.
```{r}
Avg<- sqldf("SELECT pclass, sex, Survived,
avg(age) AS AvgAge
FROM titanic
GROUP BY pclass, sex, Survived", drv='SQLite')
Avg
library(ggplot2)
ggplot(Avg, aes(x=Avg$AvgAge, y=Avg$Pclass, group=Avg$Sex)) +
  geom_line(aes(col = Avg$Survived))
```
    iv. What is the name, age, sex and pclass of the 5 oldest and 5 youngest persons who died?
```{r}
oldest<- sqldf("select name, age, sex, pclass from  titanic
where survived = 0
order by age desc limit 5 ;", drv='SQLite')
oldest
```   
    
v. On average which Passenger Class is more expensive?
```{r}
sqldf("select pclass, avg(fare) as avg_fare from titanic
              group by pclass;")
```

c. Notice the following `R` codes and explain what it is doing.
```{r}
library(RSQLite)
conn <- dbConnect(RSQLite::SQLite(), "titanicDB")
dbWriteTable(conn, name = "titanic", value = titanic, overwrite=TRUE)
dbListTables(conn)
```
1rst -> Create a connection to a DBMS (database management system), to the titanic database.
2nd -> Writes the titanic sql dataframe to a database table.
3rd -> Returns the unquoted names of titanic tables accessible through this connection.

d. Use package `dplyr` to obtain the same result as you did in 1b(iii) above. For this use the connection string `conn` and the function `tbl()`. Store the result in an object called `meanAge`.
```{r}
library(dplyr)
meanAge <- titanic %>%
  group_by(titanic$Pclass,titanic$Sex, titanic$Survived) %>% 
  summarise_at(vars(Age),
              list(name = mean))
meanAge
```

e. Show the `SQL` query to create `meanAge` in 1(d) using the function `show_query()`
```{r}
#show_query(Titanic)
```



2. **Extracting twitter data:** In this problem we would like to extract data from twitter. For this refer to the documentation in the following link.

`https://github.com/geoffjentry/twitteR/`

a. **Twitter API:** Set up twitter API using any of the following methods. Make sure you installed all the packages as mentioned in the class.

**Method 1:** Read Getting Started section of the above link and create a twitter application by going to the link `https://apps.twitter.com/`. Once you created your application connect twitter from `R` using the secrets and keys obtained from your twitter application. 


```{r, eval=FALSE}
library(twitteR)
api_key <- "your api key"  
api_secret <- "your api secret" 
access_token <- "your access tocken"
access_token_secret <- "your access tocken secret"

setup_twitter_oauth(api_key,api_secret,access_token,
                    access_token_secret)
```

**Method 2:** If you don't like creating an account with twitter and going through all the trouble, you can use my keys (ssh, don't tell anyone). For this download the `hw5-twitter-auth` file from blackboard and load it as follows.

```{r, eval=FALSE}
load("hw5-twitter-auth")

library(twitteR)
setup_twitter_oauth(api_key,api_secret,access_token,
                    access_token_secret)
```

b. Now search twitter messages for "data science job". Display few job informations.

```{r}
#dsTweet <- searchTwitter("data science job")
#print(dsTweet)
```


c. Search 300 tweets using the hash tag `#chess` and save them in an object called `rTweets`. Show the top 7 sources of tweets (such as android or iphone) in a ordered bar plot. 

```{r}
#rTweets <- searchTwitter("#chess", n=300)
#sources <- sapply(rTweets, function(x) x$getStatusSource())
#sources <- gsub("</a>", "", sources)
#sources <- strsplit(sources, ">")
#sources <- sapply(sources, function(x) ifelse(length(x) > 1, x[2], x[1]))
#source_table = table(sources)
#df <- data.frame(names(source_table),source_table)
#ggplot(df[df$Freq>8,], aes(reorder(sources,Freq),Freq)) +
#  geom_bar(stat="identity") + coord_flip()
```



d. Notice that the object `rTweets` is a list. Convert it into a data frame using function `twListToDF` and store it in an object called `dTweets`. Display some data from `dTweets`.
```{r}
#dTweets <- twListToDF(rTweets)
#head(dTweets$text)
```


e. `dTweets` has a column showing the time the tweet was created. Generate a plot showing number of tweets on each of the hours. Add a smooth line overlaid on your plot.
```{r}
#datetime<-strftime(dTweets$created, "%Y-%m-%d %H:%M:$S")
#time<-strftime(datetime, format="%H")

#df <- as.data.frame(table(time))
#ggplot(data=df, aes(x=time, y=Freq)) +
#  geom_bar(stat="identity")
```


f. Arrange the dataframe `dTweets` based on the `retweetCount`. While doing this select only columns `text, screenName, retweetCount`. Store the data in a object called `mostTweets`. Display five texts that are most retweeted.
```{r}
#mostTweets <- sqldf("select text, screenName, retweetCount from  dTweets
#order by retweetCount ;", drv='SQLite')
#mostTweets$text[(length(mostTweets$text)-5):length(mostTweets$text)]
```


g. Generate a bar chart showing top 15 screen names and count of retweets from `mostTweets`. Order the bars based on the retweet counts.
```{r}
#mostTweets <- arrange(dTweets, "retweetCount") %>% subset(select = c("text", "screenName", "retweetCount"))
#mostTweets <- mostTweets[order(mostTweets$retweetCount, decreasing = TRUE), ]
#mostTweetsTop <- mostTweets[1:15,2:3]
#mostTweetsTop
#ggplot(mostTweetsTop, aes(reorder(screenName, retweetCount),retweetCount)) + geom_bar(stat="identity") + coord_flip()
```

